---
title: "Killer Whale ZIP GLMM"
author: "Olivia Beck"
date: "`r Sys.Date()`"
output: html_document
---



## General Set Up

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(warning = F)

### Libraries ----------------------------
# data package 
library(aninet) #https://github.com/MNWeiss/aninet

# for plotting
library(ggplot2) 

# for creating design matrix
library(dplyr)
library(tidyr)

# igraph objects + functions
library(igraph) 

# parallelization is used in simulation
library(foreach) 
library(parallel) 

# model estimation
library(GLMMadaptive) 
library(MASS) 

# library(glmmTMB) Laplace model fitting algorithm
# library(centiserve) Other centrality statistics

# Other needed Functions 
source("funcs/simulation-funcs.R")# function to run simulation
source("funcs/munk-czado-funcs.R") # functions to calculate trimmed mallows distance
source("funcs/whale-plot-funcs.R") # functions to make plots

### Setting up Parallelization  ----------------------
n.cores <- parallel::detectCores() -1 

my.cluster <- parallel::makeCluster(
  n.cores, 
  type = "PSOCK"
)
my.cluster

doParallel::registerDoParallel(cl = my.cluster)

paste(("is registered:"), (foreach::getDoParRegistered()))

paste(("workers:"), (foreach::getDoParWorkers()))


```

## Section 0: Plotting the Raw Data

### Build the Surfacing Data

```{r}
### load the data ---------------------
### Bulding Whale graph
gwhale.contact <- graph_from_adjacency_matrix(srkw_contact, 
                                              mode = "undirected",
                                              weighted = T)

gwhale.surface <- graph_from_adjacency_matrix(srkw_surfacing, 
                                              mode = "undirected",
                                              weighted = T)

n <- gorder(gwhale.contact)
  
  
### clusters -------------------------
set.seed(252784)
m0 <- factor(srkw_attributes$matriline, labels = 1:6)
m1 <- cluster_walktrap(gwhale.contact, steps = 3)$membership
m2 <- cluster_leading_eigen(gwhale.contact)$membership
m3 <- cluster_infomap(gwhale.contact, E(gwhale.contact)$weight, modularity = T)$membership
m4 <- cluster_fluid_communities(gwhale.contact, 4)$membership

m5 <- cluster_walktrap(gwhale.surface, steps = 3)$membership
m6 <- cluster_leading_eigen(gwhale.surface)$membership
m7 <- cluster_infomap(gwhale.surface, E(gwhale.surface)$weight, modularity = T)$membership
#cluster_edge_closeness(gwhale.surface, weights = log(E(gwhale.surface)$weight + 1, base = 10))$membership
m8 <- cluster_fluid_communities(gwhale.surface, 4)$membership

### Node Positions for plotting 

# Define the string
position.raw <- 
  "id,age,sex,asc,matriline,X,Y
  J40,15,0,af,J14s,4.5,0.5
  J45,10,1,im,J14s,3.25,0.59
  J31,24,0,af,J11s,3.75,4.5
  J35,21,0,af,J17s,0,2.5
  J39,16,1,am,J11s,3.5,3.25
  J44,10,1,im,J17s,-1,4.5
  J53,4,0,im,J17s,0.8,3.75
  J56,0,0,im,J11s,4.5,2.5
  J47,9,1,im,J17s,-1,3.5
  J26,28,1,am,J16s,-2.6,2.25
  J37,18,0,af,J14s,2.25,0
  J38,16,1,am,J22s,-4.25,0.25
  J41,14,0,af,J19s,0,-1.5
  J46,10,0,im,J17s,0.25,5
  J49,7,1,im,J14s,3.75,-0.25
  J42,12,0,af,J16s,-4,2.25
  J22,34,0,af,J22s,-3.5,-0.5
  J27,28,1,am,J11s,5.25,3.75
  J19,40,0,af,J19s,-0.5,-0.5
  J51,4,1,im,J19s,-1.75,-1.25
  J16,47,0,af,J16s,-5,2.75
  J36,20,0,af,J16s,-3.5,3"

# Read the string into a data frame
node_df <- read.csv(textConnection(position.raw))
node_positions <- as.matrix(node_df[, c("X", "Y")])

```



### Surface Data Plots 

```{r}
size = scales::rescale(srkw_attributes$age, c(7,20))
r = range(c(E(gwhale.surface)$weight,E(gwhale.contact)$weight ))

width.surface = (E(gwhale.surface)$weight - min(E(gwhale.surface))) /( r[2] - r[1])
width.surface = width.surface * (10 - 0.5) + 0.5


# color palett for matrilines
palett0 <- c( "#9c179e", "#22a884", "#f0f921", "#fca636" ,"#414487", "#e16462"  )
p0 <- palett0[m0]

#funtion to get edge colors 
get_egdge_colors <- function(w){
  max.w <- max(w)
  prop.w <- log(w) / log(max.w)
  scaled.prop.w <- prop.w / 2 + 1/2
  edge.color <- 
    sapply(scaled.prop.w, 
           function(x){adjustcolor( "black", alpha.f = x)})
  return(edge.color)
  
}

plot(gwhale.surface,
     layout = node_positions,
     vertex.label = V(gwhale.surface)$name,
     vertex.label.cex = 1,
     vertex.label.color = "black",
     vertex.color = p0,
     vertex.size = size + 5,
     vertex.shape = ifelse(srkw_attributes$sex==0, "circle", "square"),
     vertex.frame.color = NA,
     edge.width =width.surface,
     edge.color =get_egdge_colors(E(gwhale.surface)$weight )
)


```


### Contact Data Plots 


```{r}
#edge widths for plotting
width.contact = (E(gwhale.contact)$weight - min(E(gwhale.contact))) /( r[2] - r[1])
width.contact = width.contact * (10 - 0.5) + 0.5


#assign cluster colors 
palett1 <- c("#f0f921", "#f89540", "#e16462", "#0d0887", "#9c179e" )
palett2 <- c( "#9c179e","#0d0887",   "#f89540",   "#f0f921", "#e16462")

palett3 <- c("#9c179e", "#e16462", "#f0f921" )
  #c( "#9c179e", "#fca636",   "#0d0887", "#e16462", "#6a00a8", "#f0f921")
palett4 <- c( "#f89540", "#9c179e",  "#0d0887", "#f0f921")

p1 <- palett1[m1]
p2 <- palett2[m2]
p3 <- palett3[m3]
p4 <- palett4[m4]



plot(gwhale.contact,
     layout = node_positions,
     vertex.label = V(gwhale.contact)$name,
     vertex.label.cex = 1,
     vertex.label.color = "black",
     vertex.color = p0,
     vertex.size = size + 5,
     vertex.shape = ifelse(srkw_attributes$sex==0, "circle", "square"),
     vertex.frame.color = NA,
     edge.width =width.contact,
     edge.color =get_egdge_colors(E(gwhale.contact)$weight )
)




plot(gwhale.contact, 
     layout = node_positions,
     vertex.label = NA,
     #vertex.label.color = "black",
     vertex.color = p1,
     vertex.frame.color = NA,
     main= "(A) Walktrap", 
     edge.width =width.contact,
     edge.color = "black")

plot(gwhale.contact, 
     layout = node_positions,
     vertex.label = NA,
     #vertex.label.color = "black",
     vertex.color = p2,
     vertex.frame.color = NA,
     main= "(B) Leading Eigenvalue", 
     edge.width =width.contact,
     edge.color = "black")

plot(gwhale.contact, 
     layout = node_positions,
     vertex.label = NA,
     #vertex.label.color = "black",
     vertex.color = p3,
     vertex.frame.color = NA,
     main= "(C) Infomap", 
     edge.width = width.contact,
     edge.color = "black")

plot(gwhale.contact, 
     layout = node_positions,
     vertex.label = NA,
     #vertex.label.color = "black",
     vertex.color = p4,
     vertex.frame.color = NA,
     main= "(D) Fluid Communities", 
     edge.width =width.contact,
     edge.color = "black")
```



## Section 1: Surfacing Data and Simulation

### Raw Centrality Statistics 

```{r}
### Centrality Needed for Simulation - Surfacing -----------------


whale.degree <- degree(gwhale.surface)
whale.strength <- strength(gwhale.surface)
whale.eigen <- eigen_centrality(gwhale.surface)$vector
whale.close <- igraph::closeness(gwhale.surface, normalized = T) 
whale.between <- igraph::betweenness(gwhale.surface, normalized = T)
dat.plot.true <- data.frame(degree.sim = whale.degree,
                            strength.sim = whale.strength,
                            eigen.sim = whale.eigen,
                            close.sim = whale.close ,
                            between.sim = whale.between,
                            id = 20000)

```


### Build the Design Matrix

```{r}

#### Design Matrix ----------------------------------

mother.calf <- data.frame(
  mother = c("J16", "J16", "J16", "J37", "J35", "J22", "J31", "J19", "J41"),
  calf =   c("J26", "J26", "J42", "J49", "J47", "J38", "J56", "J41", "J51")
)



#seetins up ASC2 
srkw_attributes <- 
  srkw_attributes %>% 
  dplyr::mutate(ASC2 = case_when(
    sex == 0 & age < 7 ~ "IM", #imature
    sex == 1 & age < 10 ~ "IM", #imature
    sex == 0 & age >=7 & age <38 ~ "RF", #reproductive age female 
    sex == 1 & age >=10 ~ "RM", #reproductive age male
    sex == 0 & age >=38 ~ "PF" #post reproductive age female
  )) %>% 
  dplyr::mutate(RA = case_when(
    sex == 0 & age < 7 ~ "IM", #imature
    sex == 1 & age < 10 ~ "IM", #imature
    sex == 0 & age >=7 & age <38 ~ "R", #reproductive age female 
    sex == 1 & age >=10 ~ "R", #reproductive age male
    sex == 0 & age >=38 ~ "P" #post reproductive age female
  ))




## Make Design Matrix

design.matrix <- as.data.frame(t(combn(1:n, 2)))

design.matrix[, c("name1", "name2", 
                  "kinship", "is.mother", 
                  "age1", "age2", "agediff",
                  "sex1", "sex2", "samesex", "sexgroup", "mategroup",
                  "t1", "t2", "t12", "tdiff",
                  "surface", "contact",
                  "m1", "m2",
                  "ASCi", "ASCj", "mate")] <- NA



for(i in 1:nrow(design.matrix)){
  
  #node numbers
  v1 <- design.matrix$V1[i]
  v2 <- design.matrix$V2[i]
  
  #node names
  n1 <- design.matrix$name1[i] <- V(gwhale.contact)$name[v1]
  n2 <- design.matrix$name2[i] <- V(gwhale.contact)$name[v2]
  
  #id
  id1 <- srkw_attributes$id==n1
  id2 <- srkw_attributes$id==n2
  
  
  #kinship coefficent 
  design.matrix$m1[i] <- srkw_attributes$matriline[id1]
  design.matrix$m2[i] <- srkw_attributes$matriline[id2]
  design.matrix$kinship[i] <- srkw_kinship[n1, n2]
  design.matrix$is.mother[i] <- any(apply(mother.calf, 1,
                                          function(row){all(row == c(n1, n2) | all(row == c(n2, n1)))
                                          }))
  
  
  #age coefficents
  design.matrix$age1[i] <- srkw_attributes$age[id1]
  design.matrix$age2[i] <- srkw_attributes$age[id2]
  design.matrix$agediff[i] <- abs(design.matrix$age1[i] - design.matrix$age2[i] )
  
  #sex coefficent
  design.matrix$sex1[i] <- ifelse(srkw_attributes$sex[id1] == 0, "F", "M")
  design.matrix$sex2[i] <- ifelse(srkw_attributes$sex[id2] == 0, "F", "M")
  design.matrix$samesex[i] <- design.matrix$sex1[i] == design.matrix$sex2[i]
  
  design.matrix$sexgroup[i] <- ifelse(!design.matrix$samesex[i], "N", design.matrix$sex1[i])
  
  #ASC 
  design.matrix$ASCi[i] <- srkw_attributes$RA[id1]
  design.matrix$ASCj[i] <- srkw_attributes$RA[id2]
  design.matrix$mate[i] <- 
    (srkw_attributes$ASC2[id1] == "RF" & srkw_attributes$ASC2[id2] == "RM") |
    (srkw_attributes$ASC2[id1] == "RM" & srkw_attributes$ASC2[id2] == "RF")
  
  design.matrix$mategroup[i] <- 
    case_when(
      design.matrix$mate[i] == TRUE ~ "RA", 
      design.matrix$sexgroup[i] == "F" ~ "F", 
      design.matrix$sexgroup[i] == "M" ~ "M" ,
      design.matrix$samesex[i] == F & any(srkw_attributes$RA[id1|id2] == "IM") ~ "NO", 
      design.matrix$samesex[i] == F & any(srkw_attributes$RA[id1|id2] == "P") ~ "NO", 
    )
  
  #Time of Filming 
  design.matrix$t1[i] <- srkw_sampling[n1, n1]
  design.matrix$t2[i] <- srkw_sampling[n2, n2] 
  design.matrix$t12[i] <- srkw_sampling[n1, n2] 
  design.matrix$tdiff[i] <- design.matrix$t1[i] + design.matrix$t2[i] -   design.matrix$t12[i]
  
  #response variables
  design.matrix$contact[i] <- srkw_contact[n1, n2]
  design.matrix$surface[i] <- srkw_surfacing[n1, n2]
}


design.matrix$is.mother <- as.factor(design.matrix$is.mother)
design.matrix$mategroup <- as.factor(design.matrix$mategroup)
design.matrix$samesex <- as.factor(design.matrix$samesex)


## graph summary info
sum(design.matrix$surface == 0) / nrow(design.matrix) * 100
mean(design.matrix$surface[design.matrix$surface != 0])
max(design.matrix$surface[design.matrix$surface != 0])

sum(design.matrix$contact == 0) / nrow(design.matrix) * 100
mean(design.matrix$contact[design.matrix$contact != 0])
max(design.matrix$contact[design.matrix$contact != 0])



```


### Run the Surfacing Model Selection Simulation


```{r}
### Running The Simulation ------------------------
n.models <- 2

model_results_surface <- vector(mode = "list", length = n.models)
simulation_results_surface <- vector(mode = "list", length = n.models)
mallow_results_surface <- vector(mode = "list", length = n.models)



set.seed(94353) 
for(p in 1:n.models){
  
  #define cluster membership
  membership <- m0
  
  #assign groups, group1 - group2 is the same as group2 - group1 because network is undirected
  design.matrix$group <- NA
  for(i in 1:nrow(design.matrix)){
    mm1 <- membership[srkw_attributes$id == design.matrix$name1[i]]
    mm2 <-  membership[srkw_attributes$id == design.matrix$name2[i]]
    
    # design.matrix$group[i] <- mm1 == mm2
    # design.matrix$group[i] <- (as.numeric(as.character(mm1)) * (mm1 == mm2))
    design.matrix$group[i] <- paste0(sort(c(mm1, mm2)), collapse = "")

  }
  design.matrix$group <- as.factor(design.matrix$group)
  
  
  
  #################################
  ### Modeling 
  #################################
  
  ### GLMADAPTIVE
  # ZIP with Random Effects
  # removing kinship from the zero inflated part because the hessian is uninvertible. 
  
  n_groups <- length(levels(design.matrix$group) )
  #levels(design.matrix$group) <-  LETTERS[1:10]
  #design.matrix$group
  
  if(p == 1){
    mod_zir <-
      GLMMadaptive::mixed_model(
        fixed = surface ~  mategroup*agediff+ is.mother*agediff + log(tdiff + 1), 
        random = ~ 1| as.factor(group), 
        data = design.matrix,
        family = GLMMadaptive::zi.poisson(), 
        zi_fixed = surface ~   mategroup * agediff +log(tdiff + 1) , #no is mother because no 0's in surfacing
        zi_random = ~ 1 | as.factor(group),
        max_coef_value = 10000)
  }else if(p==2){
    mod_zir <-
      GLMMadaptive::mixed_model(
        fixed = surface ~ samesex*agediff+ log(tdiff + 1), 
        random = ~ 1| as.factor(group), 
        data = design.matrix,
        family = GLMMadaptive::zi.poisson(), 
        zi_fixed = surface ~   samesex * agediff+log(tdiff + 1) , #no is mother because no 0's in surfacing
        zi_random = ~ 1 | as.factor(group),
        max_coef_value = 10000)
  }
  
  
  model_results_surface[[p]] <- mod_zir
  
  #######################################
  ### Running the simulations 
  #######################################
  
  NSIM = 10000
  
  # 3. ZIP with Random Effects 1
  simulation_results_surface[[p]]  <- sim_results(mod_zir , design.matrix, NSIM)
  mallow_results_surface[[p]] <- mallow_test(simulation_results_surface[[p]])
  
}
```

### General Model Summary Information 

#### Model 1

```{r}
summary(model_results_surface[[1]])
marginal_coefs(model_results_surface[[1]], std_errors = TRUE)
```

#### Model 2

```{r}
summary(model_results_surface[[2]])
marginal_coefs(model_results_surface[[2]], std_errors = TRUE)
```

#### AIC and BIC

```{r}
lapply(model_results_surface, BIC)
lapply(model_results_surface, AIC)
```

### Posterior Modes 

```{r}
post.modes1 <- model_results_surface[[1]]$post_modes
post.modes2 <- model_results_surface[[2]]$post_modes

post.modes <- cbind(post.modes1, post.modes2) 
colnames(post.modes) <- c("Counts1", "Zeros1", "Counts2", "Zeros2")
post.modes <- 
  post.modes %>% 
  as.data.frame() %>%
  mutate_all(round, 3)

matrilines <- c("J11s" , "J14s", "J16s", "J17s", "J19s", "J22s")
post.modes <-
  unique(t(apply(expand.grid(matrilines, matrilines), 1, sort))) %>% 
  as.data.frame() %>% 
  mutate(names = paste0(V1, "-", V2)) %>% 
  cbind(post.modes) %>% 
  dplyr::select(-c(V1, V2))

post.modes
```


### Model Validation - Counts

#### Model 1

```{r}
# From GLMMadaptive vignettes
set.seed(28397)
## Model 1
par(mar = c(2.5, 2.5, 0, 0), mgp = c(1.1, 0.5, 0), cex.axis = 0.7, cex.lab = 0.8)
y <- design.matrix$surface
y[y > 0] <- log(y[y > 0])
x_vals <- seq(min(y), max(y), length.out = 500)
out <- simulate(model_results_surface[[1]], nsim = 100, acount_MLEs_var = TRUE)
ind <- out > sqrt(.Machine$double.eps)
out[ind] <- log(out[ind])
rep_y <- apply(out, 2, function (x, x_vals) ecdf(x)(x_vals), x_vals = x_vals)
matplot(x_vals, rep_y, type = "l", lty = 1, col = "lightgrey", 
        xlab = "log(Response)", ylab = "Empirical CDF")
lines(x_vals, ecdf(y)(x_vals))
legend("bottomright", c("Simulated data           ", "Observed data      "), lty = 1, 
       col = c("lightgrey", "black"), bty = "n", cex = 0.8)

```


#### Model 2

```{r}
set.seed(783495)

### Model 2 
par(mar = c(2.5, 2.5, 0, 0), mgp = c(1.1, 0.5, 0), cex.axis = 0.7, cex.lab = 0.8)
y <- design.matrix$surface
y[y > 0] <- log(y[y > 0])
x_vals <- seq(min(y), max(y), length.out = 500)
out <- simulate(model_results_surface[[2]], nsim = 100, acount_MLEs_var = TRUE)
ind <- out > sqrt(.Machine$double.eps)
out[ind] <- log(out[ind])
rep_y <- apply(out, 2, function (x, x_vals) ecdf(x)(x_vals), x_vals = x_vals)
matplot(x_vals, rep_y, type = "l", lty = 1, col = "lightgrey", 
        xlab = "log(Response)", ylab = "Empirical CDF")
lines(x_vals, ecdf(y)(x_vals))
legend("bottomright", c("Simulated data           ", "Observed data      "), lty = 1, 
       col = c("lightgrey", "black"), bty = "n", cex = 0.8)

```


### Model Vaildation - Zeros 

#### Model 1

```{r}

###Plotting Zero Inflation factors -----------------------
# Manually predict the zero-inflation probabilities
zi_predictions <- 
  GLMMadaptive::predict(
    model_results_surface[[1]], 
    design.matrix, 
    type = "zero_part" )

# Create a data frame with predicted values and the original data
predicted_df <- data.frame(
  group = design.matrix$group,
  surface = design.matrix$surface,
  samesex = design.matrix$samesex,
  zi_prob = zi_predictions
)



# Use ggplot to visualize

## Zero Probabilities 
set.seed(9348724)
predicted_df %>% 
  mutate(is.zero = surface == 0) %>%
  filter(grepl("2", group)) %>% 
  ggplot(aes(x = group, y = zi_prob, color = is.zero)) +
  geom_jitter(width = 0.15, height = 0) +
  labs(title = "Predicted Zero-Inflation Probabilities", 
       subtitle = "Matrilineal Line J14's",
       y = "Zero-Inflation Probability", x = "Matrilineal Line") +
  scale_x_discrete(labels = c("J11s", "J14s", "J16s", "J17s", "J19s", "J22s")) +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero")) +
  theme_minimal()




### Estimated Zero Inflation Probabilities
# "outlier" of (9, 0.42) is J16-J42 which are mother-and-daughter (35 year age diff)
# "outlier" of (9, 0.12) is J19-J51 which are grandmother-and-grandson
# their physical attributes (fixed effects) suggest they may not surface together often
# but their familial relationship (random effects) suggests they do surface often
# only fixed effects are shown in this plot
predicted_df %>% 
  mutate(is.zero = surface == 0) %>% 
  ggplot(aes(x = surface, y = zi_prob)) +
  geom_point() +
  xlim(0, 25) +
  theme_minimal() +
  xlab("Observed Surfacings") +
  ylab("Zero Inflation Probabilty Estimate") 

### Maximum ZI prob over 25
predicted_df %>% 
  mutate(is.zero = surface == 0) %>% 
  filter(surface > 25) %>% 
  arrange(zi_prob) %>% 
  tail(1)


predicted_df %>% 
  mutate(is.zero = surface == 0) %>% 
  left_join(
    post.modes %>% 
      dplyr::select(names, Zeros1)%>% 
      mutate(group = factor(names, labels = levels(design.matrix$group)))
  ) %>% 
  mutate(step1 = LaplacesDemon::logit(zi_prob)) %>% 
  mutate(step2 = step1 + Zeros1) %>% 
  mutate(step3 = LaplacesDemon::invlogit(step2)) %>% 
  ggplot(aes(x = surface, y = step3)) +
  geom_point() +
  xlim(0, 25) +
  theme_minimal() +
  ggtitle("ZI Probs with Random Effects")+
  xlab("Observed Surfacings") +
  ylab("Zero Inflation Probabilty Estimate") 


#Verify if the predictions make sense
# design.matrix %>% 
#   dplyr::select(group, surface) %>% 
#   group_by(group) %>% 
#   summarise(zeros = sum(surface == 0),
#             count = n()) %>% 
#   filter(grepl("2", group))

# design.matrix %>% 
#   dplyr::select(group, surface) %>% 
#   group_by(group) %>%
#   filter(surface>0) %>%
#   summarise(nzavg = mean(surface),
#             count = n()) %>% 
#   filter(grepl("1", group))
# 
# table(srkw_attributes$matriline)
# 
# design.matrix %>% 
#   dplyr::select(group, surface) %>% 
#   filter(surface>0) %>% 
#   filter(grepl("1", group)) %>%
#   ggplot(aes(x = group, y = surface)) +
#   geom_point()
# 
# design.matrix %>% 
#   dplyr::select(group, surface)  %>% 
#   filter(grepl("1", group)) %>% 
#   ggplot(aes(x = surface)) +
#   geom_histogram() +
#   facet_wrap(~ group)
```

#### Model 2 

```{r}

###Plotting Zero Inflation factors -----------------------
# Manually predict the zero-inflation probabilities
zi_predictions <- 
  GLMMadaptive::predict(
    model_results_surface[[2]], 
    design.matrix, 
    type = "zero_part" )

# Create a data frame with predicted values and the original data
predicted_df <- data.frame(
  group = design.matrix$group,
  surface = design.matrix$surface,
  samesex = design.matrix$samesex,
  zi_prob = zi_predictions
)



# Use ggplot to visualize

## Zero Probabilities 
 set.seed(4545)
predicted_df %>% 
  mutate(is.zero = surface == 0) %>%
  filter(grepl("2", group)) %>% 
  ggplot(aes(x = group, y = zi_prob, color = is.zero)) +
  geom_jitter(width = 0.15, height = 0) +
  labs(title = "Predicted Zero-Inflation Probabilities", 
       subtitle = "Matrilineal Line J14's",
       y = "Zero-Inflation Probability", x = "Matrilineal Line") +
  scale_x_discrete(labels = c("J11s", "J14s", "J16s", "J17s", "J19s", "J22s")) +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero")) +
  theme_minimal()



### Estimated Zero Inflation Probabilities
# "outlier" of (9, 0.48) is J16-J42 which are mother-and-daughter (35 year age diff)
# "outlier" of (9, 0.19) is J19-J51 which are grandmother-and-grandson (36 year age diff)
# their physical attributes (fixed effects) suggest they may not surface together often
# but their familial relationship (random effects) suggests they do surface often
# only fixed effects are shown in this plot
predicted_df %>% 
  mutate(is.zero = surface == 0) %>% 
  ggplot(aes(x = surface, y = zi_prob)) +
  geom_point() +
  xlim(0, 25) +
  theme_minimal() +
  xlab("Observed Surfacings") +
  ylab("Zero Inflation Probabilty Estimate") 


### Maximum ZI prob over 25
predicted_df %>% 
  mutate(is.zero = surface == 0) %>% 
  filter(surface > 25) %>% 
  arrange(zi_prob) %>% 
  tail(1)

# 
# 
# #Verify if the predictions make sense
# design.matrix %>% 
#   dplyr::select(group, surface) %>% 
#   group_by(group) %>% 
#   summarise(zeros = sum(surface == 0),
#             count = n()) %>% 
#   filter(grepl("2", group))
# 
# design.matrix %>% 
#   dplyr::select(group, surface) %>% 
#   group_by(group) %>%
#   filter(surface>0) %>%
#   summarise(nzavg = mean(surface),
#             count = n()) %>% 
#   filter(grepl("1", group))
# 
# table(srkw_attributes$matriline)
# 
# design.matrix %>% 
#   dplyr::select(group, surface) %>% 
#   filter(surface>0) %>% 
#   filter(grepl("1", group)) %>%
#   ggplot(aes(x = group, y = surface)) +
#   geom_point()
# 
# design.matrix %>% 
#   dplyr::select(group, surface)  %>% 
#   filter(grepl("1", group)) %>% 
#   ggplot(aes(x = surface)) +
#   geom_histogram() +
#   facet_wrap(~ group)
```


### Subject predicted responses

For all predicted values, if $\hat{y}_{ij} < 0.5$, then $\hat{y}_{ij}$ is considered a prediced zero.  


#### Model 1 Predictions

```{r}


subject_predictions <- GLMMadaptive::predict(
    model_results_surface[[1]], 
    design.matrix, 
    type = "subject_specific" ) 


subject_predictions_df <- data.frame(
  group = design.matrix$group,
  surface = design.matrix$surface,
  samesex = design.matrix$samesex,
  pred = subject_predictions
)
# number of predicted zeros vs true zeros 
subject_predictions_df %>%
  mutate(predict.zero = pred < 0.5) %>% 
  group_by(group) %>% 
  summarise(zi_predictions = sum(predict.zero),
            zi_true = sum(surface==0),
            count = n()) %>% 
  filter(grepl("2", group))

subject_predictions_df <- 
  subject_predictions_df %>%
  mutate(predict.zero = pred < 0.5) %>%
  mutate(is.zero = surface ==0) %>%
  mutate(error = surface - pred) %>%
  mutate(rel_error = error / pred ) 

## Plot for predicted coutns 
set.seed(989)
subject_predictions_df %>% 
  filter(grepl("2", group)) %>% 
  filter(!predict.zero) %>%
  ggplot(aes(x = group, y = pred, color = is.zero)) +
  geom_jitter(width = 0.15, height = 0) +
  labs(title = "Predicted Counts - Subject Specific", 
       subtitle = "With Matrilineal Line J14's",
       y = "Counts of Surfacing", x = "Matrilineal Line") +
  scale_x_discrete(labels = c("J11s", "J14s", "J16s", "J17s", "J19s", "J22s")) +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero")) +
  theme_minimal()#+theme(panel.grid.major = element_line(color = "darkgray"))


## Error plot 
subject_predictions_df %>% 
  ggplot(aes(x = surface, y = error)) +
  geom_point() +
  facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Relative Error") +
  ggtitle("Relative Error Vs True Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero"))


```

#### Model 1 Error 

```{r}
subject_predictions_df %>% 
  filter(!is.zero) %>% 
  ggplot(aes(x = surface, y = error)) +
  geom_point() +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Error") +
  ggtitle("Error Vs True Counts Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero"))

subject_predictions_df %>% 
  filter(pred >= 0.5) %>% 
  ggplot(aes(x = surface, y = rel_error)) +
  geom_point() +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Relative Error") +
  ggtitle("Relative  of Counts Predictions Vs True Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero"))

subject_predictions_df %>% 
  filter(is.zero) %>% 
  ggplot(aes(x = surface, y = abs(error))) +
  geom_boxplot(coeff = 1000) +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Absolute Error") +
  ggtitle("Absolute Error Vs True Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero")) +
  theme_minimal()


### Error for the true counts
subject_predictions_df %>% 
  filter(!is.zero) %>%
  ggplot(aes(x = surface, y = error)) +
  geom_point() +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("Observed Counts") +
  ylab("Error") +
  ggtitle("Error vs Observed Counts") +
  xlim(0, 75) + # removes one point at (162, -11)
  theme_minimal()

# Error for the True zeros
subject_predictions_df %>% 
  filter(is.zero) %>%
  ggplot(aes( x = abs(error))) +
  geom_histogram(bins = 10) +
  #facet_wrap(~is.zero, scales = "free" )+
  ylab("Frequency") +
  xlab("Absolute Error") +
  ggtitle("Distribution of Absolute Errors for Observed Zeros") +
  theme_minimal() 
  

```



#### Model 2 Predictions

```{r}


subject_predictions <- GLMMadaptive::predict(
    model_results_surface[[2]], 
    design.matrix, 
    type = "subject_specific" ) 


subject_predictions_df <- data.frame(
  group = design.matrix$group,
  is.mother = design.matrix$is.mother,
  surface = design.matrix$surface,
  samesex = design.matrix$samesex,
  pred = subject_predictions
)
# number of predicted zeros vs true zeros 
subject_predictions_df %>%
  mutate(predict.zero = pred < 0.5) %>% 
  group_by(group) %>% 
  summarise(zi_predictions = sum(predict.zero),
            zi_true = sum(surface==0),
            count = n()) %>% 
  filter(grepl("2", group))


subject_predictions_df <- 
  subject_predictions_df %>%
  mutate(predict.zero = pred < 0.5) %>%
  mutate(is.zero = surface ==0) %>%
  mutate(error = surface - pred) %>%
  mutate(rel_error = error / pred )

set.seed(989)
subject_predictions_df %>% 
  mutate(predict.zero = pred < 0.5) %>%
  mutate(is.zero = surface ==0) %>%
  filter(grepl("2", group)) %>% 
  ggplot(aes(x = group, y = pred, color = is.zero)) +
  geom_jitter(width = 0.15, height = 0) +
  labs(title = "Predicted Response - Subject Specific", 
       subtitle = "Matrilineal Line J14's",
       y = "Counts of Surfacing", x = "Matrilineal Line") +
  scale_x_discrete(labels = c("J11s", "J14s", "J16s", "J17s", "J19s", "J22s")) +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero")) +
  theme_minimal()+
  theme(panel.grid.major = element_line(color = "darkgray"))
```


#### Model 2 Error 

```{r}
subject_predictions_df %>% 
  mutate(error = surface - pred) %>%
  mutate(rel_error = error / pred ) %>%
  mutate(is.zero = surface ==0) %>%
  filter(!is.zero) %>% 
  ggplot(aes(x = surface, y = error)) +
  geom_point() +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Error") +
  ggtitle("Error Vs True Counts Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero"))

subject_predictions_df %>% 
  mutate(error = surface - pred) %>%
  mutate(rel_error = error / pred ) %>%
  mutate(is.zero = surface ==0) %>%
  filter(pred >= 0.5) %>% 
  ggplot(aes(x = surface, y = rel_error)) +
  geom_point() +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Relative Error") +
  ggtitle("Relative  of Counts Predictions Vs True Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero"))

subject_predictions_df %>% 
  mutate(error = surface - pred) %>%
  mutate(rel_error = error / pred ) %>%
  mutate(is.zero = surface ==0) %>%
  filter(is.zero) %>% 
  ggplot(aes(x = surface, y = abs(error))) +
  geom_boxplot(coeff = 1000) +
  #facet_wrap(~is.zero, scales = "free" )+
  xlab("True Observation") +
  ylab("Absolute Error") +
  ggtitle("Absolute Error Vs True Observation") +
  scale_color_manual(values = c("#F8766D", "#619CFF"),
                     name = "True Observation",
                     labels = c("Count", "Zero")) +
  theme_minimal()



### Error for the true counts
subject_predictions_df %>% 
  filter(!is.zero) %>%
  ggplot(aes(x = surface, y = error)) +
  geom_point() +
  xlab("Observed Count") +
  ylab("Error") +
  ggtitle("Error vs Observed Counts") +
  xlim(0, 75) + # removes one point at (162, -4.3)
  theme_minimal()

# Error for the True zeros
subject_predictions_df %>% 
  filter(is.zero) %>%
  ggplot(aes( x = abs(error))) +
  geom_histogram(bins = 10) +
  ylab("Frequency") +
  xlab("Absolute Error") +
  ggtitle("Distribution of Absolute Errors for Observed Zeros") +
  theme_minimal() 

```


### ECDF Plots 

```{r}
#### ECDF Plots ----------------------

dist_plots_surface <- plot_ecdf(simulation_results_surface)

egg::ggarrange(plots = dist_plots_surface$eigen, 
               nrow = 1, ncol = n.models)

egg::ggarrange(plots = dist_plots_surface$close, 
               nrow = 1, ncol = n.models)

```



### Mallows/Wasserstein Distance Plots 

```{r}
wass_plots_testing <- plot_wass_density_surface(mallow_results_surface)
wass_plots_testing[[1]]
wass_plots_testing[[2]]
```


#### Marginal Effects Plots for Model 2

```{r}

# Interaction plots --------------------------
### Model 2
nDF2 <- with(design.matrix, 
             expand.grid(agediff = seq(min(agediff), 40, length.out = 30),
                         samesex = levels(samesex),
                         tdiff = quantile(tdiff, c(0.25, 0.5, 0.75)),
                         group = levels(group)))
nDF2$surface <- predict(model_results_surface[[2]], nDF2)  

marginal_coefs(model_results_surface[[2]], std_errors = TRUE)

#model_results_surface[[2]]$model_frames
#fitted(model_results_surface[[2]], type = "subject_specific")

nDF2 <- with(design.matrix, 
             expand.grid(agediff = seq(min(agediff), 40, length.out = 30),
                         samesex = levels(samesex),
                         tdiff =  quantile(tdiff, c(0.4, 0.6, 0.8)),
                         group = levels(group)))

# get marginal effects
nMM2 <- model.matrix(~ samesex*agediff + log(tdiff+1), data = nDF2)

marg.counts <- marginal_coefs(model_results_surface[[2]], std_errors = T)
betac <- marginal_coefs(model_results_surface[[2]])# (as.matrix(model_results_surface[[2]]$coefficients))
betac <- as.matrix(betac$betas)
head(nMM2)

nDF2$pred <- nMM2 %*% betac


sex.labs <- c("Opposite Sexes", "Same Sex")
names(sex.labs) <- c(FALSE, TRUE)

ggplot(nDF2, aes(x = agediff, color = as.factor(tdiff))) +
  geom_line(aes(y = exp(pred))) + 
  facet_wrap(~samesex,
             labeller = labeller(samesex = sex.labs)) +
  scale_color_manual(values = c("#F8766D", "#00BA38", "#619CFF"),
                     labels = c("0.4", "0.6", "0.8"),
                     name = "Quantile of Film \nCo-Occurrence") +
  xlab("Age Difference") + 
  ylab("E(Count)") +
  ggtitle("Expected Marginal Effects for the Surfacing Counts") + 
  theme_minimal() +
  theme(axis.line.x = element_line())+
  scale_y_continuous(expand=c(0,1.5)) +
  scale_x_continuous(expand=c(0,0.5)) +
  geom_vline(xintercept=0)

```




## Section 2: Contact Network - Picking the best cluster

### Setting up Needed information 

```{r}

# store centrality statistics for plotting later
whale.degree <- degree(gwhale.contact)
whale.strength <- strength(gwhale.contact)
whale.eigen <- eigen_centrality(gwhale.contact)$vector
whale.close <- closeness(gwhale.contact, normalized = T)
whale.between <- betweenness(gwhale.contact, normalized = T)
dat.plot.true <- data.frame(degree.sim = whale.degree,
                            strength.sim = whale.strength,
                            eigen.sim =whale.eigen,
                            between.sim = whale.between,
                            close.sim = whale.close,
                            id = 20000)


```


### Running the Contact network simulation 

```{r}
set.seed(43453)
model_results <- vector(mode = "list", length = 4)
simulation_results <- vector(mode = "list", length = 4)
mallow_results <- vector(mode = "list", length = 4)

set.seed(9941429) 
for(p in 1:4){

  # make design matrix 
  design.matrix <- as.data.frame(t(combn(1:n, 2)))

  #define cluster membership
  membership <- 
    case_when(
    p == 1 ~ m1,
    p == 2 ~ m2,
    p == 3 ~ m3,
    p == 4 ~ m4
  )
  
  
  design.matrix[, c("name1", "name2", 
                    "kinship", 
                    "age1", "age2", "agediff",
                    "sex1", "sex2", "samesex", 
                    "t1", "t2", "t12", "tdiff",
                    "surface", "contact",
                    "m1", "m2")] <- NA
  
  
  
  for(i in 1:nrow(design.matrix)){
    
    v1 <- design.matrix$V1[i]
    v2 <- design.matrix$V2[i]
    
    n1 <- design.matrix$name1[i] <- V(gwhale.contact)$name[v1]
    n2 <- design.matrix$name2[i] <- V(gwhale.contact)$name[v2]
    
    design.matrix$m1[i] <- membership[v1]
    design.matrix$m2[i] <- membership[v2]
    
    design.matrix$kinship[i] <- srkw_kinship[n1, n2]
    
    design.matrix$age1[i] <- srkw_attributes$age[srkw_attributes$id == n1]
    design.matrix$age2[i] <- srkw_attributes$age[srkw_attributes$id == n2]
    design.matrix$agediff[i] <- abs(design.matrix$age1[i] - design.matrix$age2[i] )
    
    design.matrix$sex1[i] <- srkw_attributes$sex[srkw_attributes$id == n1]
    design.matrix$sex2[i] <- srkw_attributes$sex[srkw_attributes$id == n2]
    design.matrix$samesex[i] <- design.matrix$sex1[i] == design.matrix$sex2[i]
    
    design.matrix$t1[i] <- srkw_sampling[n1, n1]
    design.matrix$t2[i] <- srkw_sampling[n2, n2] 
    design.matrix$t12[i] <- srkw_sampling[n1, n2] 
    design.matrix$tdiff[i] <- design.matrix$t1[i] + design.matrix$t2[i] -   design.matrix$t12[i]
    
    design.matrix$contact[i] <- srkw_contact[n1, n2]
    design.matrix$surface[i] <- srkw_surfacing[n1, n2]
  
}


#assign groups, group1 - group2 is the same as group2 - group1 because network is undirected
design.matrix$group <- NA
for(i in 1:nrow(design.matrix)){
  mm1 <- design.matrix$m1[i]
  mm2 <- design.matrix$m2[i]
  
  design.matrix$group[i] <- paste0(sort(c(mm1, mm2)), collapse = "")
  
}


design.matrix$group <- as.factor(design.matrix$group)
design.matrix$samesex <- as.factor(design.matrix$samesex)




#################################
### Modeling 
#################################

### GLMADAPTIVE
# ZIP with Random Effects
# removing kinship from the zero inflated part because the hessian is uninvertible. 

n_groups <- length(levels(design.matrix$group) )
#levels(design.matrix$group) <-  LETTERS[1:10]
#design.matrix$group


mod_zir <-
  GLMMadaptive::mixed_model(
    fixed = contact ~ agediff * samesex+ log(tdiff + 1), 
    random = ~ 1| as.factor(group), 
    data = design.matrix,
    family = GLMMadaptive::zi.poisson(), 
    zi_fixed = contact ~ agediff * samesex +  log(tdiff + 1) ,
    zi_random = ~ 1 | as.factor(group),
    max_coef_value = 10000)


model_results[[p]] <- mod_zir

#######################################
### Running the simulations 
#######################################

NSIM = 10000

# 3. ZIP with Random Effects 1
simulation_results[[p]]  <- sim_results(mod_zir , design.matrix, NSIM)
mallow_results[[p]] <- mallow_test(simulation_results[[p]])

}

```

### Model Summary

```{r}
lapply(model_results, summary)

```


### ECDF Plots 

```{r}


dist_plots <- plot_ecdf(simulation_results)
egg::ggarrange(plots = dist_plots$eigen,
               nrow = 1, ncol = 4)
egg::ggarrange(plots = dist_plots$close,
               nrow = 1, ncol = 4)

```


### Mallows/Wasserstin Distance Plots 

```{r}
wass_plots <- plot_wass_density_contact(mallow_results )
wass_plots[[1]]
wass_plots[[2]]
```


